---
title: "Homework 1 on Bass Model"
author: "Aida Martirosyan"
date: "2023-09-28"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
```

```{r, results=FALSE, echo=FALSE}
# Load necessary libraries
libs <- c('ggplot2','ggpubr','knitr','diffusion')

load_libraries<-function(libs){
  new_libs <- libs[!(libs %in% installed.packages()[,"Package"])] 
  if(length(new_libs)>0) {
    install.packages(new_libs)
    } 
  lapply(libs, library, character.only = TRUE)
}

load_libraries(libs)
```

# Introduction
\vspace{10px}
This markdown file is for the *Homework 1 of the Marketing Analytics - Fall 2023*
course. The goal of the homework is to use Bass Model, which is a technique to 
estimate Market Entry activities. For example, if someone wants to create new 
product, but has no financial resources to explore market, it is a good choice 
to use Bass Model to estimate market potential and number of adopters of the new 
product over time. The main steps of using the model includes choosing new 
innovation, then find look-alike product from the past, estimate innovation and 
imitation rates (p,q). In next steps each part of the assignment will be 
implemented and explained with visuals and texts.


# Step 1: Innovation selection
\vspace{10px}
I visited TIME magazine to choose one of the interesting innovations of the year.
As I am very interested in AI and Robotics, I found a product which attracted me
a lot. The product name is **An AI 3D Printer: Bambu Lab X1 3D Printer**. Now there
are many 3D printers in the world, but the mentioned product is different from 
existing ones as it provides magical improvements to users due to its AI powered
system. It is possible to print in sixteen various colors with this product twice 
as fast as existing printers. Another advancement is that users can control the 
AI 3D Printer with its mobile app remotely. To create the innovation it took 
approximately two years from founders, but as a result the product idea became 
very popular and successful. 


# Step 2: Look-alike Innovation selection
\vspace{10px}
The first thing that came to my mind thinking about look-alike innovation from 
past for AI 3D Printer, of course, are the usual printers that use paper to 
print and create contexts. I did research to collect information about history 
of printers. Here are some points that I underlined:
\vspace{10px}

* Printing is very old, during 3500 BC, people used cylinder seals to certify 
documents. Next comes woodblock printing, during which people used wood blocks
with specific designs to press onto paper.
* In the 19th century, Charles Babbage designed the first computer printer. 
However, this mechanical printer became available in 2000 by the Science Museum 
in London.
* In 1968 Japanese company Epson created first electronic printer, EP-101.
* In 1969 Gary Starkweather, an engineer at Xerox invented the first laser 
printer.

\vspace{10px}
I selected the **Laser Printers** as a look-alike innovation from past for AI 3D 
Printer Innovation. First laser printers, Xerox 9700 were introduced by Xerox 
company. Laser printers after short time became popular and widely
used by people as they were providing speed, high quality and high-volume results.
The printers used laser beam to print text and graphics on paper. There are some 
types of laser printers. One type includes the printers which can only print in
white and black, there are also colorful printers which can print in four colors: 
cyan, magenta, yellow, black. However, during time, as technologies evolve, emails
and e-books became highly used by people, the usage of printers declined. And here
comes my selection of the first step, AI 3D Printer, which presents 3D modeling 
that is the next stage for laser printers. Using similar technique as laser 
printers, 3D printers give chance users to create physical objects. In case of 
AI 3D Printer, it provides further advantages, including twice faster speed, 16
colors and remote control over the system, which proves that this product is
real improvement of past laser printers.


# Step 3: Look-alike Innovation Sales dataset extraction
\vspace{10px}
In this stage I tried to search some datasets of laser printers in the website
statista. After the research I found out that there are no appropriate data for
some specific types of printers or specific company's printers in the website. 
Then I decided not to focus on some specific company product but try to find 
general sales dataset of printers of some country. 

\vspace{10px}
I separated three possible choices for data. The first one was good as it included 
longer period of time series with which will be more interesting to work,
however the sales included not only printers but also scanners and multifunctional 
devices. So, considering that my target was printers I avoided this data. The 
second possible data was showing Samsung's laser printer sales, which was 
appropriate for my case, however as data included short period I also avoided this 
one. Finally, I chose the dataset that included the number of printers sold in 
Brazil between 2014 and 2020. 

The selected dataset is also good one, as it shows yearly printers sales in
million units. Of course, the data is not worldwide, but it gives local 
information about printer sales in Brazil.


# Step 4: Bass model parameters estimation
\vspace{10px}
Here is what the data looks like; the printer sales in million units are shown 
from 2014 to 2020 in Brazil:
```{r, echo=FALSE}
excel_file_path <- "printer-unit-sales-in-brazil-2014-2020.xlsx"
sheet_name <- "Data"
data <- read_excel(excel_file_path, sheet = sheet_name)

kable(data)
```

\newpage
Here is the barplot of the sales:
\vspace{10px}


```{r, echo=FALSE}
sales = ggplot(data, aes(x = Year, y = Sales)) + geom_bar(stat='identity') +
labs(title = 'Yearly printer sales in Brazil', y = 'Sales (in million units)')
print(sales)
```


\vspace{10px}
Let us use diffusion library to estimate Bass model parameters:
\vspace{10px}

```{r, echo=FALSE}
diff_m = diffusion(data$Sales)
p_estimated=round(diff_m$w,4)[1]
q_estimated=round(diff_m$w,4)[2]
m=round(diff_m$w,4)[3]
print(diff_m)
```
*Coefficient of innovation(p)* is approximately 0.05. This shows the rate by which
new users or innovators adopt the innovation. Innovators are people who are ready 
to buy the product by just knowing about the product from advertisements or 
product info.

*Coefficient of imitation(q)* is 0. This shows the rate by which imitators adopt
the innovation. In our case as it is 0, it means we do not have imitators in the
model, which is not good. Imitators are people who start buying the product by
seeing others experiments with the product.

*Market potential(m)* shows the total number of people that can buy the product, 
aka number of adopters in the market.

I am going to use the above estimated parameters during next steps...


\vspace{10px}
Modeling f(t) - the fraction of the total market that adopts at time t and F(t) 
- the fraction of the total market that has adopted up to and including time t:
\vspace{10px}

```{r, echo=FALSE}
# Define functions for f (t) and F(t):
bass.f <- function(t, p, q) {
  ((p + q)**2 / p) * exp(- (p + q) * t) / (1 + (q / p) * exp(- (p + q) * t))**2
}

bass.F <- function(t,p,q){ (1-exp(-(p+q)*t))/
    (1+(q/p)*exp(-(p+q)*t))
}
```

```{r, echo=FALSE}
 cum_ad1 = ggplot(data.frame(t = c(1, 7)), aes(t)) +
  stat_function(fun = bass.F, args = c(p=0.0523, q=0)) +
  labs(title = 'Printers - cumulative adoptions')

time_ad1 = ggplot(data.frame(t = c(1, 7)), aes(t)) +
  stat_function(fun = bass.f, args = c(p=0.0523, q=0)) +
  labs(title = 'Printers - adoptions at time t')

ggarrange(cum_ad1, time_ad1)
```

\newpage
Visualizing f(t):
\vspace{10px}

```{r, echo=FALSE}
time_ad =  ggplot(data.frame(t = c(1:7)), aes(t)) +
  stat_function(fun = bass.f, args = c(p=0.0523, q=0)) +
  labs(title = 'f(t)')
sm_sales = ggplot(data, aes(x = Year, y = Sales)) +
  geom_bar(stat = 'identity') +
  ggtitle('Printer sales, mln units')
ggarrange(time_ad, sm_sales)
```
\vspace{10px}
As we can see the estimation is not as good, yes it grasped the main trend, 
however it is not showing minor fluctuations of the barplot.

\newpage
# Step 5: Predictions of the diffusion of the innovation

Making predictions by multiplying f (t) by m:
\vspace{10px}

```{r, echo=FALSE}
data$pred_sales = bass.f(1:7, p = 0.0523, q = 0)*53.8462
ggplot(data, aes(x = Year, y = Sales)) +
  geom_bar(stat = 'identity') +
  geom_point(mapping = aes(x=Year, y=pred_sales), color = 'red')
```
As we can see from above prediction results, they are rather good. The red points 
are close to the peaks of bars. Of course, result is not perfect, but is not
very bad too.

# Step 6: Potential market share estimation using Fermi's logic


Now let us try to estimate the number of adopters for AI 3D Printer product
worldwide. I will use Fermi estimation, which is a method by which one can 
approximately estimate potential market for products. To do estimation for my 
selected product, I will consider the following facts:

* The world population is approximately 8 billion.
* Mainly the target market will include industries and professionals from fields
like design, technology and manufacturing. There are globally 507K graphic 
designers, 300 million employees in Manufacturing sector. As there was no 
specific number of employees of IT field, I will take approximately 250 million
employees. Of, course the numbers are approximated as even numbers taken from the 
Internet do not include all the employees. So, we can approximately say that our
target market is 10% of whole population.

```{r, echo=FALSE, results=FALSE}
# calculating the number of employees in manufacturing sector
3540000 + 7911000 + 4090000 + 5012000 + 4499000 + 128869000 + 10958000 + 9154000 +
10260000 + 16363000 + 612000 + 3396000 + 2332000 + 10388000 + 57244000 +
16381000 + 898000 + 3069000 
```
* We should consider that from the above mentioned professionals, only the ones
will buy the product which have the resources and need to create AI based 3D 
objects. I think from the beginning their number will not be so high, as the
prices will be high, however let us consider that 5% of them will adopt the
product.
* So final estimate for AI 3D Printer adopters (m) will be equal to 8 bilion's 10 
percent and then its 5 percent: 40 million people. 


\newpage
# Sources

* TIME magazine. *An AI 3D Printer*. (2022). Retrieved from: https://time.com/collection/best-inventions-2022/6226978/bambu-lab-x1-3d-printer/
* GSM BARCODING. *THE HISTORY OF PRINTERS *. (2020). Retrieved from:https://www.barcoding.co.uk/history-of-printers/#:~:text=The%20first%20electronic%20printer%20was,Epson%20(Shinshu%20Seiki%20Co.)
* hc. *Laser Printer*. (2023). Retrieved from: https://history-computer.com/laser-printer/
* statista. *Number of printers sold in Brazil from 2014 to 2020*. (2021). 
Retrieved from: https://www.statista.com/statistics/863044/brazil-printer-unit-sales/
* World Population Review. *World Population by Country 2023*. (2023). 
Retrieved from: https://worldpopulationreview.com
* colorlib. *Graphic Design Statistics*. (2023). 
Retrieved from: https://colorlib.com/wp/graphic-design-statistics/
* BROOKINGS. *Global manufacturing scorecard*. (2018). 
Retrieved from: 
https://www.brookings.edu/articles/global-manufacturing-scorecard-how-the-us-compares-to-18-other-nations/


